\section{Approach}
\label{sec:approach}
\subsection{Overview}

MPTA-Repair is a counterexample-guided repair workflow for
multi-property, multi-counterexample clock-bound repair of
timed-automata control models \cite{clarke2000cegar}. Its input is a faulty model
\texttt{M}, a property set \texttt{Φ\ =\ \{φ1,\ ...,\ φm\}}, and
configuration parameters such as the maximum number of refinement
rounds, candidate limits, and a global timeout. The repairable bounds
are not given as a separate user input; they are extracted from the
numerical clock bounds occurring in transition guards and location
invariants \cite{koelbl2019clock}. The output is either a repaired model
\texttt{M'} together with a set of clock-bound edits,
or a failure report explaining why no admissible full-property repair
was found within the configured budget.

The workflow is designed around two principles. First, candidate repair
generation may be driven by local diagnostic traces, but candidate
acceptance must be global. A candidate that repairs one violated
property may cause another property to fail, because the same clock,
location, transition, or synchronization may participate in several
timing requirements. MPTA-Repair therefore keeps collecting
counterexamples from newly violated properties and uses them to further
restrict the set of admissible repair candidates. This refinement
continues until a candidate satisfies the complete property set and
passes admissibility checking, or until the configured budget is
exhausted. In our evaluation, we use a fixed timeout, for example 10
minutes per repair instance, to avoid unbounded refinement \cite{jia2011mutation}.

Second, multiple properties and multiple diagnostic traces are not
treated as unrelated repair subproblems. A single faulty timing bound
may contribute to several violations, and several traces may expose
different timing realizations of the same underlying defect. MPTA-Repair
therefore maintains a pool of property-trace obligations and searches
for a clock-bound assignment that satisfies all currently encoded
obligations. Relation analysis among properties, traces, and repairable
bounds is used to reduce redundant work and guide the search, but it is
not used as a substitute for verification \cite{clarke2000cegar}.

At a high level, MPTA-Repair proceeds as follows. It verifies the model
against all properties, collects TDTs for violated properties, extracts
repairable clock-bound values, and builds a joint MaxSMT repair problem
from the current property-trace pool. The resulting assignment is
instantiated into a candidate model. The candidate is then checked
against the complete property set and against a TARTAR-style
admissibility criterion \cite{koelbl2019clock}. If validation fails, the rejected
assignment is blocked, newly observed diagnostic traces are added, and
the repair loop continues.

\subsection{Bound Variation and Joint Trace Encoding}

MPTA-Repair follows the bound-variation view used in clock-bound repair
\cite{koelbl2019clock}. Let a repairable clock-bound value be a numerical bound
\texttt{β} occurring in an atomic clock constraint \texttt{x\ ∼\ β},
where \texttt{x} is a clock and
\texttt{∼\ ∈\ \{\textless{},\ ≤,\ =,\ ≥,\ \textgreater{}\}} is fixed by
the original model. For each such bound, MPTA-Repair introduces a bound
variation variable \texttt{v}. The repaired constraint is written as:

\texttt{x\ ∼\ (β\ +\ v)}.

The current repair pipeline changes only the numerical bound. The clock
reference, comparison operator, transition structure, location
structure, synchronization labels, and reset set are not repair
variables in the main experiment. Equality constraints can be
represented in the same syntactic form when they occur in guards or
invariants, but the operator itself is not changed.

For a diagnostic trace \texttt{τ}, MPTA-Repair constructs a TARTAR-style
trace constraint system \cite{koelbl2019clock}. Let \texttt{Xτ} denote the trace-local
variables, including delay variables and clock values at trace
positions, and let \texttt{V} denote the bound variation variables. The
bound-varied trace constraint system is written as:

\texttt{Tτ\^{}bv(Xτ,\ V)}.

It contains the usual constraints for clock initialization, time elapse,
clock resets, guard satisfaction, and invariant satisfaction along the
trace, except that bounds in guards and invariants are replaced by their
varied forms \cite{koelbl2019clock}. Bound variation variables affect only the guard
and invariant constraints in which the corresponding bound occurs.

For a violated property \texttt{φ}, the local repair obligation for a
trace \texttt{τ} is:

\texttt{Obl(τ,\ φ,\ V)\ =\ (∃Xτ\ .\ Tτ\^{}bv(Xτ,\ V))\ ∧\ (∀Xτ\ .\ Tτ\^{}bv(Xτ,\ V)\ ⇒\ Φτ,φ(Xτ))}.

The first conjunct requires that the same discrete trace skeleton
remains executable after repair. This does not require the same delays
or clock valuations as in the original counterexample; it only requires
that the same sequence of discrete transitions still has at least one
feasible timing realization. The second conjunct requires that no
feasible timing realization of this trace skeleton still violates the
considered property. Thus, the local repair obligation does not block
the functional path itself. It blocks the violating timing realizations
of that path.

For a counterexample pool \texttt{C}, where each element is a
property-trace pair \texttt{(τ,\ φ)}, MPTA-Repair builds a joint hard
constraint:

\texttt{FH\ =\ ∧(τ,φ)∈C\ Obl(τ,\ φ,\ V)\ ∧\ WF(V)\ ∧\ Blocked(V)}.

Here, \texttt{WF(V)} contains well-formedness constraints such as
non-negative repaired bounds and consistency between lower and upper
timing limits when both are present. \texttt{Blocked(V)} excludes repair
assignments that have already failed full-property regression or
admissibility. To prefer small repairs, the MaxSMT soft constraints are:

\texttt{FS\ =\ ∧v∈V\ (v\ =\ 0)}.

Maximizing the satisfied soft constraints minimizes the number of
changed clock bounds \cite{sebastiani2015optimization}. When needed, additional optimization
objectives minimize the total magnitude of the nonzero variations. This
encoding generalizes single-trace clock-bound repair from one TDT to a
pool of currently relevant property-trace obligations.

If a candidate assignment \texttt{ρ} is rejected, MPTA-Repair adds a
blocking constraint that prevents the same assignment from being
returned again. For the active variation variables \texttt{Vρ}
considered in that candidate, the exact blocking constraint can be
written as:

\texttt{Blockρ(V)\ =\ ∨v∈Vρ\ v\ ≠\ ρ(v)}.

This does not claim that all similar assignments are invalid. It only
prevents the refinement loop from repeating the same failed candidate.

\subsection{Relation-Guided Search Optimization}

As refinement proceeds, the number of violated properties, diagnostic
traces, and editable clock bounds may grow \cite{clarke2000cegar}. Encoding every
possible obligation and every editable bound without organization can
make the MaxSMT problem larger and may lead to repeated solving over
redundant evidence. MPTA-Repair therefore uses relation analysis as a
search optimization. These relations guide pruning, grouping, and
ranking, but every accepted repair is still validated by full-property
regression and admissibility.

At the property layer, MPTA-Repair handles safety properties in a
practical normalized fragment. A property is normalized into an
\texttt{A{[}{]}}-style condition over location predicates and clock or
integer comparisons \cite{vogel2022patterns}. The relation analysis currently uses
conservative syntactic and formula-level rules. Two properties are
treated as equivalent when their normalized forms are identical. A
dominance relation is used only when it can be established safely. For
example, for two upper-bound properties that are identical except for
the numerical constant,

\texttt{A{[}{]}\ (Target\ ⇒\ x\ ≤\ b1)} and
\texttt{A{[}{]}\ (Target\ ⇒\ x\ ≤\ b2)},

the first property dominates the second if \texttt{b1\ ≤\ b2}. For
lower-bound properties, the direction is reversed. Similar rules apply
only when the target predicate, clock, and operator pattern match after
normalization. Shared clocks, shared target locations, and shared
templates are recorded as dependency information, but they are not by
themselves used to remove obligations.

At the counterexample layer, MPTA-Repair maintains a pool of TDTs
associated with violated properties. Exact duplicates are removed, and
traces are annotated with their involved transitions, locations, clocks,
guard bounds, invariant bounds, and source properties \cite{koelbl2019clock}.
Structural similarity, shared prefixes, and shared repair variables are
used to organize the pool and explain why certain traces should be
considered together. However, a trace is removed from the active
obligation set only when it is an exact duplicate or when a sound
subsumption check shows that its local obligation is covered by another
encoded obligation. Otherwise, the trace remains available for joint
encoding.

At the repair-bound layer, MPTA-Repair maps each property-trace
obligation to the clock-bound values that can affect it. This mapping
helps restrict the active repair variables to those appearing in or
influencing the current counterexample pool. It also helps order
candidate generation when several bounds are available. However,
relation scores do not prove correctness, and they do not force the
solver to modify a particular bound. They only reduce the search effort
by avoiding unrelated repair variables when the current counterexample
evidence gives no reason to open them.

\subsection{Optimization Objective}

The MaxSMT objective is intentionally simple \cite{sebastiani2017maxsmt}. The primary
objective is to minimize the number of changed clock bounds by
maximizing soft constraints of the form \texttt{v\ =\ 0}. Among
candidates with the same number of changed bounds, the implementation
may minimize the total magnitude of the changes. This follows the
engineering goal of producing small and inspectable repairs.

MPTA-Repair does not optimize directly for speculative notions such as
semantic risk or causal relevance. Relation analysis is used before and
around solving to select or rank repair variables and obligations. The
solver objective itself remains centered on minimality of the edit set
and, when enabled, magnitude of the edit values.

Static consistency constraints are enforced during candidate generation
\cite{demoura2008z3}. Repaired bounds must remain in the supported numerical domain.
When a lower and an upper bound jointly constrain the same clock in a
local region, the repaired values must not make the timing interval
inconsistent. Since operators and clock references are fixed in the
current repair space, generated assignments change only numerical bound
values \cite{andre2019repairing}.

\subsection{Validation, Admissibility, and Refinement}

Candidate validation is the main correctness boundary of MPTA-Repair
\cite{clarke1999model}. After applying a candidate edit set, the repaired model is
verified against every property in \texttt{Φ}. A candidate that repairs
the triggering property but violates another property is rejected. Such
a failure means that the current encoded trace pool was insufficient:
the candidate satisfied all encoded local obligations, but the repaired
model still has an execution that violates the complete specification.
MPTA-Repair therefore collects the newly observed diagnostic trace and
adds it to the pool for the next refinement round.

After full-property regression succeeds, MPTA-Repair applies the
TARTAR-style admissibility test based on functional equivalence \cite{koelbl2019clock}.
This check is needed because a clock-bound edit may satisfy safety
properties by preventing intended behavior from occurring. TARTAR
observes that timed-language equivalence is not a suitable admissibility
criterion, because a successful timing repair is expected to remove at
least some violating timed realizations \cite{koelbl2019clock}. Instead, functional
equivalence compares the untimed languages of the original and repaired
models \cite{hopcroft2001automata}. Intuitively, the repaired model should not add or
remove functional action traces, even though the timing of those traces
may change.

In implementation, this criterion is checked by constructing untimed
automata from the original and repaired timed models and comparing their
accepted untimed languages \cite{bouyer2022zone}. If the languages differ, the repair
is rejected and a separating behavior can be reported when available.
Only candidates that pass both full-property regression and
functional-equivalence admissibility are accepted.

The refinement loop therefore combines local repair generation with
global validation. Local TDT encodings generate candidate bound
assignments; full-property regression rejects candidates that do not
satisfy the complete specification; admissibility rejects candidates
that satisfy the properties by changing functional behavior; and
blocking constraints prevent repeated failed assignments. The final
result is a repair that satisfies the currently stated multi-property
specification while preserving the intended untimed behavior of the
model.
