\section{Evaluation}
\label{sec:evaluation}
\subsection{Prototype Implementation}

We implemented MPTA-Repair as a prototype repair pipeline around a
timed-automata model parser, a verification interface, a
diagnostic-trace analyzer, a MaxSMT-based repair backend, and an
admissibility checker. The prototype uses UPPAAL verification to
evaluate properties and collect timed diagnostic traces \cite{larsen1997uppaal}, and it
uses Z3 to solve the repair constraints \cite{demoura2008z3}. The implementation
produces a structured repair report for each run, including violated
properties, collected traces, selected repair variables, candidate
edits, solver status, regression results, admissibility results,
runtime, and failure reasons.

The repair frontend extracts repairable clock-bound occurrences from
transition guards and location invariants. Each repairable site records
the template, location or transition, clock, operator, original bound,
and syntactic context. The trace analyzer maps diagnostic traces to the
repair variables that occur along the corresponding execution. The
relation analyzer builds property, trace, and repair-variable relations
used for candidate ranking and decomposition. The model writer applies
candidate bound edits and produces a repaired model for validation.

The admissibility module follows the TARTAR-style idea of functional
equivalence \cite{koelbl2019clock}. It compares the untimed behavior of the original
and repaired models and rejects candidates that add or remove functional
behavior. This check is separate from full-property verification. A
candidate must pass both the complete property set and admissibility to
be considered successful.

\subsection{Baseline}

We compare MPTA-Repair with an iterative TARTAR-style baseline \cite{koelbl2020tartar}.
The baseline is intentionally stronger than a one-shot call to a
single-trace repair tool. It verifies the current model, selects a
violated property, collects a diagnostic trace, invokes a
single-property clock-bound repair procedure, and validates each
candidate against the same full-property regression and admissibility
gates used by MPTA-Repair. If a candidate fails, the baseline continues
with additional candidates or iterations until it finds a valid repair
or reaches the same time and search budgets.

This baseline is fair because it gives single-property repair access to
global validation. However, it still repairs internally from one
property or one diagnostic trace at a time. It does not jointly analyze
relations among multiple violated properties, multiple counterexamples,
and repair variables before generating candidates.

\subsection{Datasets}

We evaluate the methods on two dataset groups. The first group consists
of benchmark timed-automata models collected from UPPAAL and XTA
benchmark sources. These models provide broader structural diversity and
allow comparison with existing clock-bound repair settings. They are
useful for testing whether the method remains effective on established
benchmark examples rather than only on one family of case studies.

The second group is an industry dataset collected from practical
industrial case studies. It contains timed-automata models related to
railway alarm and communication scenarios
\cite{gonzalez2024realtimedevs,basile2021unisig}, train-gate behavior
\cite{behrmann2004tutorial}, automotive gear control \cite{lindahl1998gear},
cardiac pacing \cite{jiang2012pacemaker}, autonomous road-junction rules
\cite{belguidoum2021junction}, mechanical ventilation
\cite{cuartas2023ventilator}, and aerospace scheduling
\cite{david2010herschel}. These models are aligned with the
industry-track motivation because they contain multi-component networks,
synchronization among components, domain-level timing requirements, and
properties that are not merely direct copies of individual guards or
invariants.

For each reference model, we select time-safety properties that the
original model satisfies. Properties are derived from the model
documentation, source papers, embedded queries, and intended industrial
requirements \cite{vogel2022patterns}. We filter out trivial properties that only mirror
a single guard or invariant, and we distinguish repair specifications
from auxiliary sanity checks or reachability queries.

\subsection{Fault Injection}

Faulty instances are generated by modifying clock-bound constants in
guards and invariants. For simple mutants, one bound is changed and the
resulting model is retained if it violates at least one selected
property. For complex mutants, two or more bounds are modified. We apply
an effectiveness filter to avoid redundant multi-fault cases: a complex
mutant is retained only if the injected edits are both meaningful,
either because each single edit can independently cause a violation or
because the combined edits create a violation that neither single edit
causes alone.

The mutation process follows the standard motivation of mutation-based
evaluation: real faulty timed-automata models are rarely available in
sufficient quantity, so faults are seeded into verified reference models
\cite{jia2011mutation}. Since this paper studies clock-bound repair, the injected
faults are restricted to the same repair space used by the methods
\cite{koelbl2019clock}. Mutants that are syntactically invalid or outside the supported
model fragment are discarded.

\subsection{Research Questions and Metrics}

The evaluation is organized around five research questions.

\textbf{RQ1: Repair effectiveness.} How often can each method find a
repair that satisfies the full property set and passes admissibility?

\textbf{RQ2: Multi-property benefit.} How often does a local repair fix
the triggering property but fail another property, and does
full-property repair avoid these regressions?

\textbf{RQ3: Multi-counterexample benefit.} Does accumulating diagnostic
traces improve repair robustness compared with repairing from a single
trace?

\textbf{RQ4: Relation-guided optimization.} Do property, counterexample,
and repair-variable relations improve repair success, runtime, or repair
size?

\textbf{RQ5: Failure analysis.} Why do some cases remain unrepaired?

We report the number of models, properties, mutants, repaired cases,
success rate, runtime, refinement rounds, counterexamples used,
candidate repairs generated, changed bounds, edit magnitude, regression
failures, admissibility rejections, timeouts, and unsupported cases. For
industry cases, we additionally report model complexity metrics
such as the number of templates, locations, transitions, clocks, and
synchronization channels.

\subsection{Overall Repair Effectiveness}

Table I summarizes the overall repair effectiveness. On the benchmark
dataset, the iterative TARTAR-style baseline repairs 72\% of the cases,
while MPTA-Repair repairs 93\%. On the industry dataset, the
difference is larger: the baseline repairs 20\% of the cases, while
MPTA-Repair repairs 76\%. These results suggest that the proposed
workflow is effective not only on general timed-automata benchmarks, but
also on harder industrial models where several properties and traces
may interact.

\input{tables/overall-results}

The baseline remains competitive on simpler benchmark cases because
many faults are local and close to the original single-trace repair
assumption. In these cases, one diagnostic trace often contains enough
information to identify a useful clock-bound edit. MPTA-Repair improves
on this setting mainly through candidate blocking, full-property
regression, and counterexample accumulation.

The industry results show a different pattern. Industrial case-study models
contain interacting components, synchronization, observer templates, and
domain-level properties. A repair that is locally valid for one trace is
frequently rejected by full-property regression or admissibility.
MPTA-Repair is more robust in this setting because it accumulates
evidence from multiple properties and traces before committing to a
repair.

\subsection{Benefit of Full-Property Regression}

Full-property regression rejects candidates that repair one property
while breaking another. This failure mode appears frequently in
multi-property models. For example, tightening an upper-bound guard may
eliminate a late-response trace, but it may also prevent a required
transition from occurring in another operating phase. Similarly,
relaxing a timeout bound may satisfy a liveness-like response monitor
while violating a safety monitor that constrains maximum waiting time.

The evaluation confirms that single-property success is not sufficient
evidence of repair validity. Several baseline candidates satisfy the
triggering property but fail another property in the same property file.
MPTA-Repair avoids accepting such candidates because every candidate is
checked against the conjunction of all selected properties. This
supports the central design decision of treating property restoration as
a global objective rather than as a sequence of independent local
obligations.

\subsection{Benefit of Multi-Counterexample Repair}

Counterexample accumulation improves robustness by reducing overfitting
to one diagnostic trace. In several cases, the first repair candidate
removed the reported TDT but left another trace feasible. After the new
trace was added to the counterexample pool, MPTA-Repair found a
different bound edit or a small joint edit set that addressed both
traces. This behavior is especially important in models with several
operational modes, where different counterexamples expose different
timing regions of the same underlying defect.

The number of refinement rounds remained modest in most successful
cases. Successful repairs usually required one to three rounds, while
harder cases accumulated more traces before either finding a repair or
timing out. The main cost of additional counterexamples is larger SMT
and regression-verification effort, but the benefit is fewer locally
overfitted repairs.

\subsection{Relation-Guided Optimization}

Relation-guided search improves both candidate quality and repair
efficiency. Property relations help avoid repeatedly solving essentially
equivalent obligations. Counterexample relations identify traces that
share path fragments or repair variables. Repair-variable relations
focus the solver on bounds that are close to the observed violations,
rather than on all syntactically editable bounds in the model.

The ablation study shows that removing relation guidance increases the
number of candidates explored and leads to more regression failures.
Removing counterexample accumulation has a larger impact on
industry cases than on benchmark cases, because industrial models
more often produce multiple related violations. Removing admissibility
increases apparent property success but accepts repairs that remove or
add functional behavior, which are not valid under our repair criterion.

\subsection{Repair Size and Runtime}

Most successful repairs are small. The median repair changes one or two
clock bounds, and the average edit magnitude remains close to the
injected fault size. This is desirable for engineering use: a repair
suggestion that changes a small number of meaningful timing thresholds
is easier to inspect than a large set of unrelated edits.

Runtime is mainly affected by the number of active traces, the number of
clocks appearing along those traces, and the number of repair variables
opened for solving. This is consistent with prior observations in
TDT-based repair: longer traces and larger constraint systems increase
quantifier elimination and solving cost \cite{koelbl2019clock}. Industry models
tend to be harder because they contain more templates, synchronizations,
and property interactions, which increase both verification cost and
admissibility-checking cost.

\subsection{Failure Analysis}

Not all cases are repairable within the current scope. We classify
failures into five categories. First, some instances have no feasible
candidate within the clock-bound repair space; the true repair may
require changing a reset, synchronization label, data update, or model
structure. Second, some candidates satisfy encoded traces but fail
full-property regression. Third, some candidates satisfy all properties
but fail admissibility because they remove or add untimed behavior.
Fourth, some instances time out during solving, verification, or
admissibility checking. Fifth, some models use unsupported syntax or
property forms that fall outside the current implementation.

This failure analysis is useful for interpreting the results.
MPTA-Repair is not a general timed-automata repair system; it is a
clock-bound repair workflow. Failures outside this repair space should
not be interpreted as failures of the central idea, but as evidence that
richer repair actions and stronger trace generation will be needed for
broader deployment.
